### 2017-11-09 paper/sections/1_introduction.tex
paper: clarify Mantle in intro, add metric dicscussion, start hacking on methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear keyspace access patterns and hot
spots that offer significant opportunity for optimization. We use the Mantle
policy engine to dynamically explore a variety of techniques, ranging from
basic algorithms and heuristics to statistical models, calculus, and machine
learning. While Mantle was originally designed for distributed file systems, we
show how it effectively decomposes the problem into manageable policies for a
different domain and service (in this case, cache management).  Our exploration
of this space results in a two policy scheme that achieves 96\% efficiency
while using only 7.6\% of the memory resources required by the base case.

@@

In this paper we present a detailed analysis of how the ParSplice application
accesses key-value pairs over the course of a long running simulation across a
variety of initial conditions.  This type of analysis (1) shows the capacity
and resource requirements of a single node and (2) will help inform our load
balancing policies for when we switch to a distributed key-value store back-end
to store EOM minima. We need to know when and how to partition the keyspace: a
smaller cache hurts performance because key-value pairs need to be retrieved
from other nodes while a larger cache has higher memory pressure.

For example, Figure~\ref{fig:motivation-regimes} shows the
cache activity of a ParSplice node and demonstrates that small changes to the
rates at which new atoms enter the simulation (\(\Delta\)) can have a strong
effect on the timing and frequency with which new EOM minima are discovered and
referenced.  The ParSplice caches on each node are trimmed when memory pressure
reaches a threshold but the number of unique keys accessed (black points
along bottom of both graphs) in Figure~\ref{fig:motivation-regimes} suggests
that a small cache of EOM minima should be sufficient.

% What is Mantle
To explore the effects of different cache management strategies, we link the
Mantle policy engine into ParSplice.  Mantle~\cite{sevilla:sc15-mantle} has
already proven to be a critical control plane for improving file system
metadata load balancing  and in this work we show its usefulness in cache
management for the changing key-value workloads generated by ParSplice.
Developers write policies for ``when" they want data moved, ``where" they want
data moved, and ``how much" of the data to move and the framework executes
these policies whenever a decision needs to be made.  This abstraction helps
developers unfamiliar with the domain quickly reason about, develop, and deploy
new policies that control temporal and spatial locality. We show that Mantle:

@@

This last contribution is explored in Section~\S\ref{sec:mantle-brains}, where
we try a range of policies from different disciplines; but more importantly, in
Section~\S\ref{sec:relation-to-file-systems}, we conclude that the collection
of policies we designed for ParSplice's cache management are very similar to
the policies in the Ceph file system that are used to load balance metadata,
suggesting that there is potential for automatically adapting and generating
policies dynamically.

>>>>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear keyspace access patterns and hot
spots that offer significant opportunity for optimization. We use the Mantle
language and policy engine to dynamically explore a variety of techniques,
ranging from basic algorithms and heuristics to statistical models, calculus,
and machine learning. While Mantle was originally designed for distributed file
systems, we show how the collection of abstractions effectively decomposes the
problem into manageable policies for a different domain and service (in this
case, cache management).  Our exploration of this space results in a two policy
scheme that achieves 96\% efficiency while using only 7.6\% of the memory
resources required by the base case.

@@

In this paper we present a detailed analysis of how the ParSplice application
accesses key-value pairs over the course of a long running simulation across a
variety of initial conditions.  This type of analysis (1) shows the capacity
and resource requirements of a single node and (2) will help inform our load
balancing policies for when we switch to a distributed key-value store back-end
to store EOM minima.  For example, Figure~\ref{fig:motivation-regimes} shows
the cache activity of a ParSplice node and demonstrates that small changes to
the rates at which new atoms enter the simulation (\(\Delta\)) can have a
strong effect on the timing and frequency with which new EOM minima are
discovered and referenced.  The ParSplice caches on each node are trimmed when
memory pressure reaches a threshold but the number of unique keys accessed
(black points along bottom of both graphs) in
Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
should be sufficient.

% What is Mantle
We use the Mantle language and policy engine to dynamically explore the effects
of different cache management strategies for the changing key-value workloads
generated by ParSplice.  Mantle~\cite{sevilla:sc15-mantle} has already proven
to be a critical control plane for improving file system metadata load
balancing  and in this work we show that the collection of abstractions
provided by Mantle is useful for reasoning about and designing different cache
management strategies.  Developers write policies for ``when" they want data
moved and ``how much" of the data to move, then the framework executes these
policies whenever a decision needs to be made.  These abstraction help
developers unfamiliar with the domain quickly reason about, develop, and deploy
new policies that control temporal and spatial locality. We show that Mantle:

@@

This last contribution is explored in
Sections~\S\ref{sec:cache-management-using-system-archiecture-knowledge}
and~\S\ref{sec:cache-management-using-domain-specific-knowledge}, where we try
a range of policies from different disciplines; but more importantly, in
Section~\S\ref{sec:relation-to-file-systems}, we conclude that the collection
of policies we designed for ParSplice's cache management are very similar to
the policies in the Ceph file system that are used to load balance metadata,
suggesting that there is potential for automatically adapting and generating
policies dynamically.

### 2017-11-09 paper/sections/2_parsplice-keyspace-analysis.tex
paper: clarify Mantle in intro, add metric dicscussion, start hacking on methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
cores per node.  Each node has 128GB of RAM and our goal is to limit the size
of the cache to 3\% of RAM\footnote{Empirically, this is a threshold that we
find to work well for most applications}. Note that this is an addition to the
30GB that ParSplice uses to manage other ranks on the same node.  A single Cray
node produced trajectories that are \(5\times\) times longer than our 10 node
CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As
a result, it reaches different job phases faster and gives us a more
comprehensive view of the workload. The performance gains compared to the
commodity clusters have more to do with memory/PCI bandwidth than network.

>>>>

All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
cores per node.  Each node has 128GB of RAM and our goal is to limit the size
of the cache to 3\% of RAM\footnote{Empirically, we find this threshold works
well for most applications}. Note that this is an addition to the 30GB that
ParSplice uses to manage other ranks on the same node.  A single Cray node
produced trajectories that are \(5\times\) times longer than our 10 node
CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As a
result, it reaches different job phases faster and gives us a more
comprehensive view of the workload. The performance gains compared to the
commodity clusters have more to do with memory/PCI bandwidth than network.

### 2017-11-09 paper/sections/3_methodology.tex
paper: clarify Mantle in intro, add metric dicscussion, start hacking on methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{itemize}
  \item motivation: Mochi load balancer microservice
  \item background: CephFS implementation
  \item library architecture, callbacks, environment
\end{itemize}

>>>>

%  \item motivation: Mochi load balancer microservice

%  \item background: CephFS implementation

%  \item library architecture, callbacks, environment
We extract Mantle as a library that is linked into a service ({\it e.g.}, file
system, key-value store, application), where administrators program policies
that are embedded into the runtime.  The service needs to be modified to (1)
provide environment of metrics (2) identify where policies are set. At runtime,
Mantle executes the admin's policies for when/where/how much and returns a
decision.

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/mantle.png}\\

  \caption{Extracting Mantle as library.\label{fig:mantle}}

\end{figure}

\begin{table}
  \centering
  \begin{tabular}{ r | l | l }
  Metrics     & Data Structure & Description \\\hline
  Cluster     & \{server \(\rightarrow\) \{metric \(\rightarrow\) val\}\}
              & resource util. for servers \\
  Time Series & [(ts, val), ..., (ts, val)]
              & accesses by timestamp (ts) \\
  && \\
              & Service      & Example \\\hline
  Cluster     & File Systems & CPU util., Inode RDs \\
              & ParSplice    & CPU util., Cache Size \\
  Time Series & File Systems & Accesses to directory \\
              & ParSplice    & Accesses to key in DB\\
  \end{tabular}
  \caption{Types of metrics exposed by the service to the policy engine using Mantle.\label{table:metrics}}
\end{table}

%of timestamp (ts), value pairs representing accesses over time types of
%metrics: key value pairs
Services expose two types of metrics with Mantle: cluster metrics for
describing resource utilization and time series metrics for describing accesses
over time. Table~\ref{table:metrics} shows how these metrics are accessed from
the policies written by administrators. Cluster metrics are read from a
dictionary indexed by server and metric name, where server is a node identifier
({e.g.}, MPI Rank, metadata server name), metric is a resource name and val is
the current reading for that metric.  Examples of \texttt{(metric, value)}
pairs are CPU or memory pressure, as shown in Figure~\ref{fig:mantle}.  For
time series metrics, the service passes a pointer to an array of
\texttt{(timestamp, value)} pairs. We use a pointer so we can pass a large
number of values, like accesses over time to a database or directory in the
file system namespace, but this limits the time series metrics to only include
values from the {\it current } node.

% policies themselves
Administrators write policies as callbacks that use the metrics from above.
Figure~\ref{fig:mantle} shows an example policy for the ``when()" callback, where the
current server (\texttt{whoami}) migrates load if it is has load
(\texttt{>0.1}) and if its neighbor server (\texttt{whoami + 1}) does not have
load (\texttt{<0.1}).
